---
title: FOSS For Spectroscopy
author: Bryan A. Hanson, DePauw University
date: "`r Sys.time()`"
output:
  html_document:
    theme: united
---

<!-- See https://kryogenix.org/code/browser/sorttable/ -->
<script src="sorttable.js"></script>

The following table collects information about free and open source software ([FOSS](https://en.wikipedia.org/wiki/Free_and_open-source_software)) for spectroscopy.  Unless otherwise noted, the software mentioned here can be used with one or more of the following instruments: NMR, IR, Raman, ESR/EPR, fluorescence and UV-Vis.  Mass *spectrometry* software is not included here.  All information is taken from the respective websites and/or repositories.  Some softwares have been described in publications.  If so, a link is provided but keep in mind the document may be behind a paywall.

Software mentioned here appears to be regularly maintained, as evidenced by one of the following:

* a commit to a repository within the past year,
* a web site updated within the past year, or
* successful continuous integration.

Additions or corrections?  Please contact [hanson@depauw.edu](mailto:hanson@depauw.edu) or submit a pull request at the [repo](https://github.com/bryanhanson/FOSS4Spectroscopy).

[![Project Status: WIP – Initial development is in progress, but there has not yet been a stable, usable release suitable for the public.](https://www.repostatus.org/badges/latest/wip.svg)](https://www.repostatus.org/#wip)
[![Build Status](https://travis-ci.com/bryanhanson/FOSS4Spectroscopy.svg?branch=master)](https://travis-ci.com/bryanhanson/FOSS4Spectroscopy)

Abbreviations:

* __EDA:__ Exploratory data analysis (unsupervised chemometrics)

<hr>

*Click on a header to sort the table*

```{r setup, echo = FALSE}
suppressPackageStartupMessages(library("knitr"))
suppressPackageStartupMessages(library("kableExtra"))
```

```{r inputValues, echo = FALSE}
# This section contains the information about the software

# Contribution Rules: One item per line, never change the order!
# Add new entries to the end of each vector
# Add a comment + line number every 5 entries
# You must add an entry to each vector, if the information is not available please add an empty string: ""

# The name of the software package or resource
name <- c(
  "ChemoSpec",
  "ChemoSpec2D",
  "speaq",
  "OpenVnmrJ",
  "nmrGlue", # 5
  "hyperSpec",
  "readJDX",
  "hyperChemoBridge",
  "nmrPipe"
  )

website <- c(
    "https://bryanhanson.github.io/ChemoSpec/",
    "https://bryanhanson.github.io/ChemoSpec2D/",
    "",
    "https://openvnmrj.org",
    "http://www.nmrglue.com/", # 5
    "http://hyperspec.r-forge.r-project.org/",
    "https://github.com/bryanhanson/readJDX",
    "",
    ""
    )

repos <- c(
  "https://github.com/bryanhanson/ChemoSpec",
  "https://github.com/bryanhanson/ChemoSpec2D",
  "https://github.com/OpenVnmrJ/",
  "",
  "https://github.com/jjhelmus/nmrglue", # 5
  "https://r-forge.r-project.org/scm/?group_id=366",
  "",
  "",
  ""
  )

pub <- c(
  "",
  "",
  "",
  "",
  "https://link.springer.com/article/10.1007%2Fs10858-013-9718-x", # 5
  "",
  "",
  "",
  ""
  )

# The language the software requires to run

lang <- c(
  "R",
  "R",
  "R",
  "?",
  "Python", # 5
  "R",
  "R",
  "R",
  "?"
  )

# The main capabilities of the software

descrip <- c(
  "EDA of 1D spectra",
  "EDA of 2D NMR",
  "Data processing & conversion between formats",
  "Data acquisition & processing (Varian instruments)",
  "Data processing, analysis, visualization & conversion", # 5
  "Hyperspectral data sets (spectra accompanied by spatial, time or concentration information)",
  "Import JCAMP-DX files to ChemoSpec/Spectra format",
  "Interconversion of hyperSpec and ChemoSpec formats",
  "Item 9"
  )

# The operating systems that the software will run on

os <- c(
  "Any running R",
  "Any running R",
  "Any running R",
  "Linux",
  "Any running Python", # 5
  "Any running R",
  "Any running R",
  "Any running R",
  "?"
  )

# Any notes

notes <- c(
  "",
  "2D NMR only",
  "",
  "NMR only",
  "nD NMR only", # 5
  "",
  "",
  "",
  ""
  )


```

<!-- If you are just adding a new resource, you don't need to edit below here. -->

```{r processValues, echo = FALSE}
# Additional processing of the input values
# Combine name and website to create hyperlink, incl pub if available
namelink <- paste("[", name, "](", website, ")", sep = "")
for (i in 1:length(namelink)) if (pub[i] != "") {
  namelink[i] <- paste(namelink[i], " ([pub](", pub[i], "))", sep = "")
}
```

```{r checkStatus, echo = FALSE}
# Automatically check if the software is apparently kept up-to-date

status <- rep(NA_character_, length(name))

```


```{r createTable, results = "asis", echo = FALSE}
DF <- data.frame(namelink, descrip, lang, os, notes, status)
names(DF) <- c("name", "description", "language", "OS", "notes", "status")
kable(DF, table.attr = "class=\"sortable\"") %>%
  kable_styling(c("striped", "bordered", "sortable"))
```
